---
title: "Gender Inequality in India"
---

## Male Against Female LFPR



```{r, eval=TRUE,echo=FALSE}

library(shiny)
library(ggplot2)
suppressMessages(library(dplyr))

# Load dataset
data <- read.csv("/Users/jaren/Desktop/NM2207 Updated Data.csv")

# Create ggplot2 line plot
ggplot(data, aes(x = X)) +
  geom_line(aes(y = Male.LFPR, color = "Male Labor Force Participation Rate")) + 
  geom_line(aes(y = Female.LFPR, color = "Female Labor Force Participation Rate")) +
  labs(
    title = "Trends in Labor Force Participation Rates in India",
    x = "Year",
    y = "Labor Force Participation Rate (%)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
    )

```

<iframe height="500" width="100%" frameborder="no" src="https://deludedjaren.shinyapps.io/Shiny2207/"> </iframe>

## Male vs Female Youth Exclusion from Education, Employment or Training

```{r, eval=TRUE,echo=FALSE}

# Load data
dataII <- read.csv("/Users/jaren/Desktop/NM2207 Updated Data II.csv")

# Create ggplot2 line plot
ggplot(dataII, aes(x = Year)) +
  geom_line(aes(y = MaleYouthNotInEET, color = "Male Youth not in Education, Employment or Training")) + 
  geom_line(aes(y = FemaleYouthNotInEET, color = "Female Youth not in Education, Employment or Training")) +
  labs(
    title = "Proportion of Youth not in Education, Employment or Training",
    x = "Year",
    y = "Youth not in Education, Employment or Training (%)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
    )


```
## Male vs Female Workers Education Level

```{r, eval=TRUE,echo=FALSE}

# Load data
dataIII <- read.csv("/Users/jaren/Desktop/NM2207 Updated Data III.csv")

# Create ggplot2 line plot
ggplot(dataIII, aes(x = Year)) +
  geom_line(aes(y = MaleLabourForceWithAdvancedEdu, color = "Male - Advanced Education")) +
  geom_line(aes(y = MaleLabourForceWithIntermediateEdu, color = "Male - Intermediate Education")) +
  geom_line(aes(y = MaleLabourForceWithBasicEdu, color = "Male - Basic Education")) +
  geom_line(aes(y = FemaleLabourForceWithAdvancedEdu, color = "Female - Advanced Education")) + 
  geom_line(aes(y = FemaleLabourForceWithIntermediateEdu, color = "Female - Intermediate Education")) +
  geom_line(aes(y = FemaleLabourForceWithBasicEdu, color = "Female - Basic Education")) +
  geom_point(aes(y = MaleLabourForceWithAdvancedEdu, color = "Male - Advanced Education")) +
  geom_point(aes(y = MaleLabourForceWithIntermediateEdu, color = "Male - Intermediate Education")) +
  geom_point(aes(y = MaleLabourForceWithBasicEdu, color = "Male - Basic Education")) +
  geom_point(aes(y = FemaleLabourForceWithAdvancedEdu, color = "Female - Advanced Education")) +
  geom_point(aes(y = FemaleLabourForceWithIntermediateEdu, color = "Female - Intermediate Education")) +
  geom_point(aes(y = FemaleLabourForceWithBasicEdu, color = "Female - Basic Education")) +
 scale_color_manual(values = c("pink", "red", "purple", "darkgreen", "darkblue", "darkgrey")) +
  labs(
    title = "Comparison of Education Levels within Labour Force",
    x = "Year",
    y = "Proportion of Labour Force (%)",
    color = "Education Level"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )


```
### Analysis of Gender Inequality in India
It can be seen from these visualization that there is evidence of gender inequality in India. The data shows that males have an unfair advantage in educational and economic opportunities. Next, we will be analysing whether or not gender inequality is worsening over time.