---
title: "W12 Submission"
author: "Jaren Ong"
date: "`r Sys.Date()`"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week 9

### What is the topic that you have finalized?

This project aims to assess gender inequality in India by comparing female and male labor force participation rates and education levels, shedding light on disparities in economic and educational opportunities.

### What are the data sources that you have curated so far?

I have attached the data set I curated in the csv file below.

```{r, eval=TRUE,echo=TRUE}

# Read data from CSV file

data <- read.csv("/Users/jaren/Desktop/NM2207 Updated Data.csv")

head(data)  # View the first few rows

summary(data)

```

## Week 10

### What is the question that you are going to answer? (Answer: One sentence that ends with a question mark that could act like the title of your data story),

Is gender equality getting worse in India?

### Why is this an important question? (Answer: 3 sentences, each of which has some evidence, e.g., "According to the United Nations..." to justify why the question you have chosen is important),

The question of whether gender equality is getting worse in India is significant due to several compelling reasons. According to the United Nations' Human Development Index, India's Gender Inequality Index (GII) ranks it as one of the countries with persistent gender disparities. India's GII, according to the 2020 Human Development Report, stood at 0.501, indicating substantial gender inequality. Furthermore, evidence from the National Family Health Survey-4 (NFHS-4) shows that despite improvements in some areas, India still grapples with gender-based issues like maternal mortality, child marriage, and violence against women, underscoring the need to assess whether the situation is improving or deteriorating. Lastly, tracking changes in gender equality in India is vital as it can inform policy and program development to address disparities and promote a more inclusive and equitable society.

### Which rows and columns of the dataset will be used to answer this question? (Answer: Actual names of the variables in the dataset that you plan to use).

I currently intend to utilize all my variables to create a story on how gender inequality is evolving in India.

## Week 11

### 1. List the visualizations that you are going to use in your project (Answer: What are the variables that you are going to plot? How will it answer your larger question?),

The main variables that I will likely focus on will be labor force participation rate for females and males.

### 2. How do you plan to make it interactive? (Answer: features of ggplot2/shiny/markdown do you plan to use to make the story interactive) ,

I used ggplot2 to create some visualizations on how female and male LFPR have altered over time.

I want to use Shiny, although I am not very sure how to do so.

### 3. What concepts incorporated in your project were taught in the course and which ones were self-learnt? (Answer: Create a table with topics in one column and Weeks in the other to indicate which concept taught in which week is being used. Leave the entry of the Week column empty for self-learnt concepts)

I will update this as my website develops.

### Include the challenges and errors that you faced and how you overcame them.

I struggle to incorporate Shiny into my website.

## Week 12

### Include the challenges and errors that you faced and how you overcame them.

Over W12, I attempted to plot ggplot2 graphs but struggled to do so for some variables due to the presence of NA values. I was not confident in tinying data in R before plotting it out. I sought help from Prof NV which allowed me to make progress as I was able to plot the graphs I wanted to answer the questions I had in mind.

That being said, given that this is the first time I am creating a website, I remained unsure on how to design it in a way that looks aesthetic. I also struggled to come up with a coherent story which conveyed the main issue that my website is trying to address. I am still in the process of coming up with questions which would allow me to answer the issue of how gender inequality has evolved over time in India.

## Final Submission

### 1. What is the theme of your data story?

### 2. Why is it important to address this question?

### 3. Why do you think the data sources that you have curated can help you answer the question?

### 4. What are the insights from the data and how are they depicted in plots?

### 5. How did you implement this entire project? Were there any new concepts that you learnt to implement some aspects of it?
